\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2014-01-28 correct.tex
updated text

\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for any
  given $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
\label{th:1}
\end{theorem}
\begin{proof}
  We use case analysis and contradiction to prove the above theorem.
  \begin{compactenum}[\hspace{0.25cm} 1.]
  \item Proof for case $T = WCRT$: Assume \mbox{$d \times WCRT < M$},
    then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition, hence,
    $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption to
    hold, $d < \frac{M}{BCRT}$, which contradicts
    Lemma~\ref{lemma:1}. The assumption $d \times WCRT >N$ is trivially
    proven from Lemma~\ref{lemma:1}.
  \item Proof for case $T = BCRT$: Like above, the assumption $d \times
    BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For $d
    \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
    $BCRT \leq WCRT$, by definition, hence, $\frac{N}{BCRT} \geq
    \frac{N}{WCRT}$, which again contradicts Lemma~\ref{lemma:1}.
  \item Proof for case $BCRT < T < WCRT$: For $d \times T < M$ to hold,
    $d < \frac{M}{T}$ should hold. By definition, $T > BCRT$, thus,
    $\frac{M}{BCRT} < \frac{M}{T}$, hence, $d < {M}{BCRT}$ should hold,
    which contradicts Lemma~\ref{lemma:1}. We can prove the other case
    $d \times T \leq N$, using the upper bound from Lemma~\ref{lemma:1}
    similarly.
  \end{compactenum}
\end{proof}

>>>>

\begin{theorem}
  Given \texttt{wait\_inbetween (M..N)} Algorithm~\ref{alg:1} gives a
  value \texttt{d}, provided $D$ is a non empty set, such that for any
	given $\hj{T}$ \hj{in the range of} $BCRT \leq T \leq WCRT$, $M \leq T \times d \leq N$ holds.
\label{th:1}
\end{theorem}
\begin{proof}
  We use case analysis and contradiction to prove the above theorem.
  \begin{compactenum}[\hspace{0.25cm} 1.]
  \item Proof for case $T = WCRT$: Assume \mbox{$d \times WCRT < M$},
    then $d < \frac{M}{WCRT}$. $BCRT \leq WCRT$, by definition, hence,
    $\frac{M}{WCRT} \leq \frac{M}{BCRT}$. Thus for the assumption to
    hold, $d < \frac{M}{BCRT}$, which contradicts
    Lemma~\ref{lemma:1}. The assumption $d \times WCRT >N$ is trivially
    proven from Lemma~\ref{lemma:1}.
  \item Proof for case $T = BCRT$: Like above, the assumption $d \times
    BCRT < M$ is trivially proven from Lemma~\ref{lemma:1}. For $d
    \times BCRT > N$ to hold, $d > \frac{N}{BCRT}$ should hold. But,
    $BCRT \leq WCRT$, by definition, hence, $\frac{N}{BCRT} \geq
    \frac{N}{WCRT}$, which again contradicts Lemma~\ref{lemma:1}.
  \item Proof for case $BCRT < T < WCRT$: For $d \times T < M$ to hold,
    $d < \frac{M}{T}$ should hold. By definition, $T > BCRT$, thus,
    $\frac{M}{BCRT} < \frac{M}{T}$, hence, $d < {M}{BCRT}$ should hold,
    which contradicts Lemma~\ref{lemma:1}. We can prove the other case
    $d \times T \leq N$, using the upper bound from Lemma~\ref{lemma:1}
    similarly.
  \end{compactenum}
\end{proof}

### 2014-01-28 delays.tex
updated text

We introduce a single \textit{derived} statement called
\mbox{\texttt{wait\_inbetween (M..N)}}, built from kernel constructs, in
the SystemJ language for real-time control. The other two constructs
\mbox{\texttt{wait\_atleast (M)}} and \mbox{\texttt{wait\_exactly (M)}}
are variants of this single derived construct.

@@

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the wait construct; $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after waiting for \textit{at least} $M$ times
  units.
\item Given a SystemJ program \texttt{wait\_exactly (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  waiting time is exact.
\end{enumerate*}

@@

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{wait\_inbetween} construct. We
first divide $M$ and $N$ with the BCRT and WCRT, respectively. This
division gives us the number of individual logical ticks required to
postpone the CD (or synchronous program) by the real-time
specification. We always $ceil$ when dividing $M$ and $floor$ when
dividing $N$ to make sure that the resultant values are integers (in
domain $\mathbb{N}^{>0}$) and these functions guarantee that the
resultant logical ticks result in real-time postponements between the
required range $[M,N]$. Next, a function $F$ maps these calculated
values to a set of equidistant integer points (values) separated by a
unit value -- these points represent all the logical ticks running at
the WCRT and the BCRT, respectively that satisfy the real-time wait
requirements. The intersection of these two sets gives all the logical
ticks that satisfy the real-time requirements invariant of the logical
time and its elasticity.

Let us revisit the HRTCS example (cf. Figure~\ref{fig:1}) to illustrate
the algorithm. From Figure~\ref{fig:1} we know that $M$ is 50.3 ms and
$N$ is 200.3 ms, respectively. Let the WCRT and BCRT be: 0.112 ms and
0.0334 ms, respectively. Thus, the algorithm proceeds as follows:

@@

\paragraph{The \texttt{wait\_exactly(M)} construct}
\label{sec:extend-techn-vari}

The \texttt{wait\_exactly(M)} variant is more interesting. Like before,
we find sets $S_1$ and $S_2$, and find the intersection of the two sets
to get the value of \texttt{d}. It is possible % (and often likely, as
% suggested by our experiments in
% Section~\ref{sec:experimental-results})
that the resultant set $D$ is empty (also possible in the case of
\texttt{wait\_inbetween(M..N)}, but never possible in case of
\texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
sets and will always contain at minimum a single element. For such cases
we provide two solutions listed below.

@@

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exactly(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD at periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exactly(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

>>>>

We introduce a single \textit{derived} statement called
\mbox{\texttt{wait\_inbetween (M..N)}}, built from kernel constructs, in
the SystemJ language for real-time control. The other two constructs
\mbox{\texttt{wait\_atleast (M)}} and \mbox{\texttt{wait\_exact (M)}}
are variants of this single derived construct.

@@

We also introduce two variants of the derived statement
\texttt{wait\_inbetween}.
\begin{enumerate*}
\item Given a SystemJ program \texttt{wait\_atleast (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau < \infty$. It is important to
  note that the lower bound of the wait construct; $M$ is \textit{not}
  an exact delay, but rather the control is allowed to proceed to the
  next statement anytime after waiting for \textit{at least} $M$ times
  units.
\item Given a SystemJ program \texttt{wait\_exact (M); p}, where $M
  \in \mathbb{Q}^{>0}$, statement $p$ is executed after real-time delay
  of $\tau$ units, such that, $M \leq \tau \leq M$. In this variant, the
  waiting time is exact.
\end{enumerate*}

@@

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively, of the \texttt{wait\_inbetween} construct. We
first divide $M$ and $N$ with the BCRT and WCRT, respectively. This
division gives us the number of individual logical ticks required to
postpone the CD (or synchronous program) by the real-time
specification. We always $ceil$ when dividing $M$ and $floor$ when
dividing $N$ to make sure that the resultant values are integers (in
domain $\mathbb{N}^{>0}$) and these functions guarantee that the
resultant logical ticks result in real-time postponements between the
required range $[M,N]$. Next, a function $F$ maps these calculated
values to a set of equidistant integer points (values) separated by a
unit value -- these points represent all the logical ticks running at
the WCRT and the BCRT, respectively that satisfy the real-time wait
requirements. The intersection of these two sets gives all the logical
ticks that satisfy the real-time requirements invariant of the logical
time and its elasticity.

Let us revisit the HRTCS example (cf. Figure~\ref{fig:1}) to illustrate
the algorithm. From Figure~\ref{fig:1} we know that $M$ is 50.3 ms and
$N$ is 200.3 ms, respectively. \hj{Timing analysis of the program has
found} WCRT and BCRT be: 0.112 ms and 0.0334 ms, respectively. Thus, the
algorithm proceeds as follows:

@@

\paragraph{The \texttt{wait\_exact(M)} construct}
\label{sec:extend-techn-vari}

The \texttt{wait\_exact(M)} variant is more interesting. Like before,
we find sets $S_1$ and $S_2$, and find the intersection of the two sets
to get the value of \texttt{d}. It is possible % (and often likely, as
% suggested by our experiments in
% Section~\ref{sec:experimental-results})
that the resultant set $D$ is empty (also possible in the case of
\texttt{wait\_inbetween(M..N)}, but never possible in case of
\texttt{wait\_atleast(M)}). Note that $S_1$ and $S_2$ can never be empty
sets and will always contain at minimum a single element. For such cases
we provide two solutions listed below.

@@

The aforementioned technique maintains the essence of synchronous and
GALS programs -- the logical tick, but in doing so is unable to provide
the exact real-time wait of $M$ time units in case of
\texttt{wait\_exact(M)} statement, thereby violating its
semantics. The second option is to execute the CD periodically at the
WCRT. WCRT determines the fastest execution speed of any CD with the
guarantee that no input events from the environment will be missed
(cf. Section~\ref{sec:mapping-logical-time}). Thus, if we simply execute
the CD at periodically at the WCRT, we can choose the value of
\textit{d} from set $S1$. This technique allows meeting the synchronous
hypothesis, while at the same time guaranteeing the semantics of the
\texttt{wait\_exact(M)} statement. But, compared to the previous
option, this technique is very inflexible and results in reduced
response time and increased power consumption.

### 2014-01-28 exps.tex
updated text

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time waits and program sizes are shown in
Table~\ref{fig:benchlist}.  {\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exactly(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exactly(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exactly(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exactly(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exactly(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

@@

HRTCS is the human response system described in
Section~\ref{sec:motivating-example}. Motor is a stepper motor
controller from~\cite{Bourke2009a}, used for producing monochrome images
on paper by correct application of current to a print head (a row of
resistors). In this example, duration of the current applied is
controlled by the feedback logic using real-time waits, which are
expressed using reactive constructs. Printer is an example
from~\cite{Schneider:1999:CRT:555233}, which models a printer and a
spooler in timed CSP. Its operation can be described as follows: as soon
as the spooler receives a print job it passes it to the printer through
a channel where actual printing process take place. This example along
with HRTCS demonstrates the use of non exact wait statements.

@@

We use the \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007} to carry out all our experiments. JOP
implements the JVM specification in hardware directly and has been shown
to be real-time amenable. We statically estimate the WCRT and BCRT of
these benchmark applications using the real-time analysis tools provided
by the JOP design framework. The \textit{Safety Critical Java} (SCJ)
specification we use for real-time analysis avoids garbage collection
(GC) and hence, GC overheads and unpredictability are completely
avoided.

@@

Table~\ref{fig:comparison} shows the result of our algorithm applied to
each benchmark program. Note that we allowed the compiler to relax the
upper real-time bounds for \texttt{wait\_exactly} and
\texttt{wait\_inbetween} when it failed to find a feasible solution.
However, a user can prohibit such behavior using compiler switch as
described in Section~\ref{sec:tool-chain-flow}.

@@

% The upper bound \texttt{N} needed to be relaxed for all examples other
% than HRTCS to obtain a feasible \texttt{d}. Therefore,
On the other hand, Algorithm \ref{alg:2} is applied to those examples,
where relaxation was required, to determine the minimum upper real-time
bound \emph{N} such that there exists a solution for \emph{D} in
Algorithm \ref{alg:1}. This algorithm gives a unique number for logical
tick delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) =
Max(\emph{$S_1$}). The compiler, therefore, will choose \emph{d} based
on the newly computed \emph{N} for the examples: Motor and AECS.

>>>>

We have carried out a number of experiments on a set of applications
with real-time constraints. The characteristics of the benchmark set
including required real-time waits and program sizes are shown in
Table~\ref{fig:benchlist}.  {\renewcommand{\arraystretch}{0.9}
%\renewcommand{\texttt}[1] {\begin{lstlisting}[language=Java,morekeywords={delay}] #1 \end{lstlisting}}
\begin{table}[h!]
	\scriptsize
	\centering
	\caption{An overview of the benchmark programs}
	\label{fig:benchlist}
	\scalebox{0.9}{
		\begin{tabular}{llp{0.5cm}p{1cm}}
\toprule
\emph{Benchmark}       & \emph{Delay constructs used in each CD} & \emph{\# of CDs} & 
\emph{Lines of code}        
\\ \midrule
HRTCS                  & \texttt{wait\_inbetween(50.3..200.3 ms)} ($CD1$) & 2                  & 56
\\ \midrule
\multirow{5}{*}{Motor} & \texttt{wait\_exact(2.4 ms)}        & \multirow{5}{*}{1} & \multirow{5}{*}{116} \\ 
											 &
                                                                                         \texttt{wait\_exact(1.667
                                                                                           ms)}      &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exact(0.05
                                                                                           ms)}       &                    &                      \\ 
											 &
                                                                                    \texttt{wait\_exact(0.3
                                                                                           ms)}        &                    &                      \\ 
											 &
                                                                                         \texttt{wait\_exact(0.733
                                                                                           ms)}      &                    &
\\ \midrule
\multirow{3}{*}{Printer}
& \texttt{wait\_inbetween(2.7..10.6 ms)} ($CD1$)      & \multirow{3}{*}{2}        & \multirow{3}{*}{39} \\
& \texttt{wait\_inbetween(1.2..30.3 ms)} ($CD2$)		  &														&											\\
& \texttt{wait\_inbetween(5.6..100.2 ms)} ($CD2$)     &														&											
\\ \midrule
\multirow{2}{*}{AECS} & \texttt{wait\_atleast(10000 ms)} ($CD1$) 
& \multirow{2}{*}{2} & \multirow{2}{*}{267} \\
& \texttt{wait\_exact(10000 ms)} ($CD2$) & &
\\ \bottomrule
\end{tabular}
}
\end{table}
}
{\renewcommand{\arraystretch}{0.6}
\begin{table*}[t!]
\centering
	\caption{Actual delays obtained for wait constructs in the
benchmark programs based on their BCRT and WCRT}
\begin{tabular}{ c c l l l l l l }
	\toprule
	\multicolumn{2}{c}{} & \emph{HRTCS} & \emph{Motor} & \emph{Printer/CD1} &
	\emph{Printer/CD2} & \emph{AECS/CD1} &	\emph{AECS/CD2} \\ 
	\midrule
	\multicolumn{2}{c}{\emph{BCRT}}	& 0.0.0396 ms	& 0.0425 ms & 0.0190 ms &
	0.0263 ms & 0.1784 ms & 0.1284 ms\\ 
	\cmidrule(r){1-8}
	\multicolumn{2}{c}{\emph{WCRT}} &0.1152 ms & 0.1096 ms & 0.0647 ms
	& 0.0491 ms & 0.3476 ms & 0.2991 ms\\ 
	\cmidrule(r){1-8}

@@

\hj{
HRTCS is the human response system described in
Section~\ref{sec:motivating-example}. }

\hj{
Motor is a stepper motor
controller from~\cite{Bourke2009a}, used for producing monochrome images
on paper by correct application of current to a print head (a row of
resistors). In this example, duration of the current applied is
controlled by the feedback logic using real-time waits, which are
expressed using reactive constructs. }

\hj{
Printer is an example
from~\cite{Schneider:1999:CRT:555233}, which models a printer and a
spooler in timed CSP. Its operation can be described as follows: as soon
as the spooler receives a print job it passes it to the printer through
a channel where actual printing process take place. This example along
with HRTCS demonstrates the use of non exact wait statements.
}

@@

We use the \textit{Java Optimized Processor}
(JOP)~\cite{jop:jnl:jsa2007}, \hj{which has been shown to be real-time
amenable}, to carry out all our experiments. We statically estimate the
WCRT and BCRT of these benchmark applications using the real-time
analysis tools provided by the JOP design framework. The \textit{Safety
Critical Java} (SCJ) specification we use for real-time analysis avoids
garbage collection (GC) and hence, GC overheads and unpredictability are
completely avoided.

@@

Table~\ref{fig:comparison} shows the result of our algorithm applied to
each benchmark program. Note that we allowed the compiler to relax the
upper real-time bounds for \texttt{wait\_exact} and
\texttt{wait\_inbetween} when it failed to find a feasible solution.
However, a user can prohibit such behavior using compiler switch as
described in Section~\ref{sec:tool-chain-flow}.

@@

% The upper bound \texttt{N} needed to be relaxed for all examples other
% than HRTCS to obtain a feasible \texttt{d}. Therefore,
On the other hand, Algorithm \ref{alg:2} is applied to those examples
where relaxation was required to determine the minimum upper real-time
bound \emph{N} such that there exists a solution for \emph{D} in
Algorithm \ref{alg:1}. This algorithm gives a unique number for logical
tick delay \emph{d} such that \emph{d} = Min(\emph{$S_2$}) =
Max(\emph{$S_1$}). The compiler, therefore, will choose \emph{d} based
on the newly computed \emph{N} for the examples: Motor and AECS.

### 2014-01-28 motivation.tex
updated text

\subsection{Motivation via examples of real-time systems programmed in a
reactive language -- SystemJ}
\label{sec:motivating-example}

@@

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,suspend}]
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(TRAP); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume logical time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; /* do something with S */}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a machine used to test human response time and collect this
data for future research purposes. The machine switches on a green light
on a touch screen for anywhere \textit{in between} 50.3 ms to 200.3 ms,
if the user can touch this green light then she is successful and an
integer \texttt{1} is sent to a database, if unsuccessful a \texttt{0}
is sent instead. Different statistics and analyses on human
responsiveness can be based on the collected data.

The SystemJ pseudo-code~\cite{amal10} for such a system, along with the
expected timing behavior of the \texttt{wait\_inbetween} statement, is
shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper,
first of all, because being a \textit{Globally Asynchronous Locally
  Synchronous} (GALS) language it helps us easily describe a larger
class of systems (like HRTCS in Figure~\ref{fig:1}) compared to purely
synchronous languages like Esterel. Moreover, being a super-set of
Esterel, SystemJ allows us to explore the marriage of real-time and
logical time, which is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of 50.3
ms to 200.3 ms a positive response is sent to the second
clock-domain. All the SystemJ programming constructs used to implement
this system are described in~\cite{amal10} and Table~\ref{tab:1}. The
\texttt{wait\_inbetween} specification, which models the non exact delay
between 50.3 ms and 200.3 ms is not supported in the current language
and cannot be compiled with the current SystemJ compiler. We introduce
such \texttt{wait\_inbetween} statements in SystemJ. Such non exact
real-time mechanisms are useful in developing systems where the
real-time delay is not known or should not be known a priori.

@@

\subsubsection{Programming periodic tasks using the exact real-time
  construct; the \texttt{wait\_exactly} statement}
\label{sec:progr-using-exact}

The previous wait statements are non exact, i.e., the execution of the
statement following the wait statement cannot be controlled exactly. The
final variant we introduce allows doing exactly that. We call it
\texttt{wait\_exactly} and this statement can be used to program exact
timeouts, periodic tasks, etc.

\begin{figure}[t!]
  \centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{pp:a}Periodically emitting signal S}
        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exactly}]
while(true) { 
 wait_exactly (1 ms); 
 emit S; 
 //do something 
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{p1:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_exactly} statement. The Y-axis is the
    time elapsed.}
  \includegraphics[scale=0.7]{FF3}
\end{SubFloat}
  \caption{Programming a periodic task with wait\_exactly}
  \label{fig:p1}
  \vspace{-10pt}
\end{figure}

One such example; periodic emission of a signal to the environment is
shown in Figure~\ref{fig:p1}. As before upon execution of
\texttt{wait\_exactly (1 ms)} statement, the time starts elapsing. But,
unlike in previous cases the following statement needs to be executed
exactly after 1 ms have passed, as shown in Figure~\ref{p1:b}.

>>>>

\subsection{Motivation via examples of real-time systems programmed in a
reactive language SystemJ}
\label{sec:motivating-example}

@@

\begin{figure}[t!]
	\vspace{-10pt}
        \begin{SubFloat}{\label{delaycode:a}A human responsiveness
            system (HRTCS)}
        \begin{lstlisting}[style=sysj,morekeywords={sustain,send,receive,abort,await,emit,present,trap,pause,exit,wait_inbetween,wait_exact,suspend}]
{ // Clock-domain 1
while(true) {
  trap(T){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;} //reaction R1
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     wait_inbetween (50.3 .. 200.3 ms); 
     exit(T); 
    } // reaction R2
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  wait_exact(1 s); // Wait for 1 second
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; 
      if(#S == 1) counter++; else counter--;}
}
\end{lstlisting}
\end{SubFloat}
\begin{SubFloat}{\label{delaycode:b}Timing diagram -- the X-axis shows
    the execution of the \texttt{wait\_inbetween} statement. The Y-axis
    is the time elapsed.}
  % \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.7]{FF1}
  % \end{figure}
\end{SubFloat}
\caption{Programming a human responsiveness system (HRTCS) with
  \texttt{wait\_inbetween}}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a machine used to test human response time and collect this
data for future research purposes. The machine switches on a green light
on a touch screen for anywhere \textit{in between} 50.3 ms to 200.3 ms,
if the user can touch this green light then she is successful and \hj{an 
integer \texttt{1} is sent to other process in order to increase an
integer counter, if unsuccessful a \texttt{0} is sent instead to
decrease the counter}. Different statistics and analyses on human
responsiveness can be based on the collected data.

The SystemJ pseudo-code~\cite{amal10} for such a system, along with the
expected timing behavior of the \texttt{wait\_inbetween} statement, is
shown in Figure~\ref{fig:1}. We use the SystemJ language in this paper, 
because being a \textit{Globally Asynchronous Locally
  Synchronous} (GALS) language it helps us easily describe a larger
class of systems (like HRTCS in Figure~\ref{fig:1}) compared to purely
synchronous languages like Esterel. Moreover, being a super-set of
Esterel, SystemJ allows us to explore the marriage of real-time and
logical time, which is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two \hj{mutually asynchronous}
clock-domains, the first clock-domain continuously displays a green
light (\texttt{GREEN\_LIGHT} signal) on the touch screen sensor and
waits for the user to respond. If the user is able to touch the screen
within the specified time of 50.3 ms to 200.3 ms a positive response is
sent to the second clock-domain. All the SystemJ programming constructs
used to implement this system are described in Table~\ref{tab:1} \hj{and
more details can be found in~\cite{amal10}}. The
\texttt{wait\_inbetween} specification, which models the non exact
delay between 50.3 ms and 200.3 ms is not supported in the current
language and cannot be compiled with the current SystemJ compiler. We
introduce such \texttt{wait\_inbetween} statements in SystemJ. Such
non exact real-time mechanisms are useful in developing systems where
the real-time delay is not known or should not be known a priori.

@@

\subsubsection{Programming periodic tasks using the exact real-time
  construct; the \texttt{wait\_exact} statement}
\label{sec:progr-using-exact}

The previous wait statements are non exact, i.e., the \hj{time of}
execution of the statement following the wait statement cannot be
controlled exactly. The final variant we introduce allows doing exactly
that. We call it \texttt{wait\_exact} and this statement can be used to
program exact timeouts, periodic tasks, etc.

\begin{figure}[t!]
  \centering
	\vspace{-10pt}
        \begin{SubFloat}{\label{pp:a}Periodically emitting signal S}
        \begin{lstlisting}[style=sysj,morekeywords={emit,trap,pause,exit,wait_exact}]
while(true) { 
 wait_exact (1 ms); 
 emit S; 
 //do something 
}
\end{lstlisting}
\end{SubFloat}

\begin{SubFloat}{\label{p1:b}Timing diagram -- the X-axis shows the
    execution of the \texttt{wait\_exact} statement. The Y-axis is the
    time elapsed.}
  \includegraphics[scale=0.7]{FF3}
\end{SubFloat}
  \caption{Programming a periodic task with wait\_exact}
  \label{fig:p1}
  \vspace{-10pt}
\end{figure}

One such example; periodic emission of a signal to the environment is
shown in Figure~\ref{fig:p1}. As before upon execution of
\texttt{wait\_exact (1 ms)} statement, the time starts elapsing. But,
unlike in previous cases the following statement needs to be executed
exactly after 1 ms have passed, as shown in Figure~\ref{p1:b}.
